%----------------------------------------
\documentclass[12pt,english]{article}
%----------------------------------------
%

%---------------------------------------------------------
\input{../packages.tex}
\input{../macros.tex}
%---------------------------------------------------------

\newcommand{\GS}{\mbox{\boldmath\textbf{GSp}}}


%-------------------------------------------
\begin{document}
%-------------------------------------------

\title{Optimization algorithms with approximation}
\author{}
\maketitle
\tableofcontents

%\begin{abstract}
%\end{abstract}
%\linenumbers
%
%==========================================
\section{Introduction}\label{sec:}
%==========================================
%
%
We consider a Hilbert space $(X,\scp{\cdot}{\cdot})$ with induced norm $\norm{\cdot}$ and the minimization of a smooth $\mu$-strictly convex function $f:X\to\R$:
%
\begin{align*}
\inf_{x\in X} f(x) = \inf\SetDef{f(x)}{x\in X}.
\end{align*}
%
We suppose that a unique minimizer $x^*$ exists.
%

Our purpose is to analyse gradient algorithms on a sequence of subspaces (finite element spaces for the PDE)
%
\begin{align*}
X_0 \subset \cdots \subset X_k \subset X_{k+1}\subset \cdots \subset X,\quad P_k: X\to X_k,
\end{align*}
%
such that a typical iteration reads: 
%
\begin{yellow}
\begin{equation}\label{eq:pgm}
x_{k+1} = x_k - t_k P_{k}\nabla f(x_k),
\end{equation}
\end{yellow}
%
where $P_k$ is the orthogonal projector on $X_k$ and $\nabla f(x)\in X$ is defined by the Riesz map. In order to generate the subspaces $X_k$, we suppose to have 
an error estimators $\eta_k:X_k\to\R$ and a refinement algorithm satisfying typical hypothesis from the theory of AFEM.
% 
In the case $f\in \mathcal S^{1,1}_{\mu,L}(X)$, $X$ finite-dimensional, and $t_k=2/(\mu+L)$ for all $k$ we have the following {convergence estimate}  (Theorem 2.1.15 in \cite{Nesterov18}) for the gradient method (\textbf{GM}):
%
\begin{equation}\label{eq:gmlipsmoothest}
\norm{x_n-x^*} \le \rho^n\norm{x_0-x^*},\quad \rho = 1-1/\kappa,
\end{equation}
%
such that $\eps>0$ is achieved in $n(\eps) = O(\kappa)\ln(1/\eps)$ iterations.
It is well-known that \textbf{GM} is not optimal for this class of functions. The accelerated gradient method (\textbf{AGM}) \cite{Nesterov18} yields an improved estimate  $n(\eps) = O(\sqrt{\kappa})\ln(1/\eps)$.

Our aim is to establish a similar iteration count for the method on a sequence of subspaces.
There is important progress of adaptive finite element methods (AFEM) for nonlinear elliptic equations, 
see \cite{ErnVohralik13a,HeidWihler20,HeidPraetoriusWihler21,HaberlPraetoriusSchimanko21,GantnerHaberlPraetorius21,HeidWihler21,HeidStammWihler21,HeidWihler22}, and our development is based on these works. However, here, we wish to 
work out the optimization point of view.
%
%\cite{}
%==========================================
\section{Notation}\label{sec:}
%==========================================
%
We throughout suppose that $f:X\to\R$ is convex and $C^1$ and we use the the Fréchet-Riesz theorem to define
%
\begin{align*}
\scp{\nabla f(y)}{x} = f'(y)(x)\quad \forall x,y\in X.
\end{align*}
%
It is then easy to see, that for all $y$ in a closed subspace $Y\subset X$ and $P_Y:X\to Y$ its orthogonal projector
%
\begin{equation}\label{eq:gradrest}
P_Y\nabla f(y) = \nabla \Rest{f}{Y}.
\end{equation}
%

%==========================================
\section{Assumptions on subspace selection}\label{sec:}
%==========================================
%

Let $X_0\subset X$ be a subspace. We suppose to have a lattice of admissible closed subspaces 
%
\begin{equation}\label{eq:}
\mathcal X(X_0) = \Set{X_0\subset Y\subset X}.
\end{equation}
%
The partial order on $\mathcal X(X_0)$ is given by $Y\ge Z$ if and only if $Y$ is a superspace of $Z$. We 
then have the finest common coarsening $Y\land Z$ and the coarsest common refinement $Y\lor Z$, respectively.
We let
%
\begin{align*}
\mathcal X(Y) = \SetDef{Z\in \mathcal X(X_0)}{Y\land Z = Y},\quad Y\in \mathcal X(x_0).
\end{align*}
%
We make the following assumptions. First we have a reliable error estimator $\eta$ such that for all $Y\in \mathcal X(X_0)$
%
\begin{align}
\label{eq:hyp:estimator:reliability}\tag{H1}
\norm{(I-P_Y)\nabla f(y)}\le& \Crel \eta(y,Y)&\quad&\forall y\in Y\\
\label{hyp:estimator:stability}\tag{H2}
\abs{\eta(y, Y)-\eta(z, Y)}\le& \Cstab \norm{y-z}&\quad&\forall y,z\in Y
\end{align}
%
and a subspace generator $Y^+=\GS(Y, y)$ such that with $0\le\qred<1$
%
\begin{align}
\label{hyp:estimator:reduction}\tag{H2}
%
\eta^2(y, Y^+)\le& \qred \eta^2(y,Y) &\quad&\forall y\in Y
%
\end{align}
%


%==========================================
\section{Gradient method with constant step-size}\label{sec:}
%==========================================
%
Here we suppose in addition that $f$ has a $L$-Lipschitz continuous gradient
%
\begin{align*}
\norm{\nabla f(x)-\nabla f(y)} \le L \norm{x-y}\quad \forall x,y\in X.
\end{align*}
%
Setting $\beta=0$ in the following algorithm, we have the standard gradient method with fixed step-size.
%
%---------------------------------------
\begin{yellow}
\begin{algorithm}[H]
\caption{GM with constant step-size} 
\label{algorithm:Descent} 
%
Inputs: $X_0, x_0\in X_0$, $t_0>0, 0\le\beta<1$, $\lambda>0$. Set $x_{-1}=x_0$ and $k=0$.
%
\begin{itemize}
\item[(1)] $y_{k} = (1+\beta)x_k - \beta x_{k-1}$.
\item[(2)] $x_{k+1} = y_k - \frac{1}{L}P_{X_k}\nabla f(y_k)$.
\item[(3)] If $\eta(x_{k+1},X_k) \ge \qred \eta(x_{k},X_k) + \lambda(f(x_k)-f(x_{k+1})$:\\ 
$\qquad X_{k+1} = \GS(X_K, x_{k+1})$,\\
Else: $X_{k+1} = X_k$.
\item[(4)] Increment $k$ and go to (1).
\end{itemize}
%
\end{algorithm}
\end{yellow}
%---------------------------------------

We will write for brevity $P_k = P_{X_k}$ etc.
By the Lipschitz-condition and convexity we have for any $x\in X$ with (\ref{eq:gradrest})
%
\begin{align*}
f(x_{k+1}) \le& f(y_k) + \scp{\nabla f(y_k)}{x_{k+1}-y_k} + \frac{L}{2}\norm{P_k\nabla f(y_k)}^2\\
=&  f(y_k) + \scp{P_k \nabla f(y_k)}{x_{k+1}-y_k} + \frac{L}{2}\norm{P_k\nabla f(y_k)}^2\\
\le& f(x) + \scp{\nabla f(y_k)}{y_k-x}-\frac{\mu}{2}\norm{x-y_k}^2 - \frac{1}{2L}\norm{P_k\nabla f(y_k)}^2
\end{align*}
%
Let $\theta = \frac{1-\beta}{1+\beta}$. Taking the last inequality $\theta$-times with $x=x^*$ and 
$1-\theta$-times with $x=x_k$, we have, setting $\Delta f_k:=f(x_k)-f(x^*)$
%
\begin{align*}
\Delta f_{k+1} - (1-\theta)\Delta f_k \le& \scp{\nabla f(y_k)}{y_k-(1-\theta) x_k -\theta x^*} - \frac{1}{2L}\norm{P_k\nabla f(y_k)}^2\\&-\frac{\mu\theta}{2}\norm{x^*-y_k}^2-\frac{\mu(1-\theta)}{2}\norm{x_k-y_k}^2
\end{align*}
%
Let
%
\begin{align*}
R_k := \theta\scp{(I-P_k)\nabla f(y_k)}{y_k-x^*},
\end{align*}
%
such that
%
\begin{align*}
\scp{\nabla f(y_k)}{y_k-(1-\theta) x_k -\theta x^*} = R_k + \scp{P_k\nabla f(y_k)}{y_k-(1-\theta) x_k -\theta x^*}
\end{align*}
%

and
%
\begin{align*}
z_k := \frac{x_k}{\theta} -\frac{1-\theta}{\theta}x_{k-1}= x_k +\frac{1-\theta}{\theta}(x_k-x_{k-1}).
\end{align*}
%
We also have with $\theta(1+\beta) = 1-\beta$
%
\begin{align*}
z_k = y_k  + \frac{1-\theta-\theta\beta}{\theta\beta}(y_k-x_{k})= y_k  + \frac{y_k-x_{k}}{\theta}
\end{align*}
%

Then with $2ab-a^2=b^2-(a-b)^2$
%
\begin{align*}
&\scp{P_k\nabla f(y_k)}{y_k-(1-\theta) x_k -\theta x^*}-\frac{1}{2L}\norm{P_k\nabla f(y_k)}^2=\\
&\frac{L}{2}\left(\norm{y_k-(1-\theta) x_k -\theta x^*}^2-\norm{x_{k+1}-(1-\theta) x_k -\theta x^*}^2\right)=\\
&\frac{\theta^2L}{2}\left(\norm{x_k+\frac{y_k-x_k}{\theta} - x^*}^2-\norm{z_{k+1}- x^*}^2\right)=
\frac{\theta^2L}{2}\left(\norm{z_k -(y_k-x_k) - x^*}^2-\norm{z_{k+1}- x^*}^2\right)
\end{align*}
%
Since with $-2ab= (a-b)^2-a^2-b^2$
%
\begin{align*}
\norm{z_k -(y_k-x_k) - x^*}^2 =& \norm{z_k - x^*}^2 - 2\theta\scp{z_k -x^*}{z_k-y_k} + \norm{y_k-x_k}^2\\
=& (1-\theta) \norm{z_k - x^*}^2 + (\theta^2-\theta)\norm{z_k-y_k}^2 + \theta \norm{y_k-x^*}^2
\end{align*}
%
we have
%
%
\begin{align*}
\Delta f_{k+1} + (1-\theta)\Delta f_k \le& \frac{\theta^2L}{2}\left((1-\theta)\norm{z_k - x^*}^2-\norm{z_{k+1}- x^*}^2\right)
+ \left(\frac{\theta^3L}{2}-\frac{\theta\mu}{2}\right)\norm{y_k-x^*}^2 + R_k
\end{align*}
%
We have
%
\begin{align*}
(1-\theta)\norm{z_k - x^*}^2 + \theta \norm{y_k-x^*}^2 =& \norm{(1-\theta)z_k+\theta y_k - x^*}^2 + \theta(1-\theta)\norm{z_k -y_k}^2\\
=& \norm{y_k  + (1-\theta)\frac{y_k-x_{k}}{\theta} - x^*}^2 + \frac{1-\theta}{\theta}\norm{y_k -x_k}^2
\end{align*}
%

%
\begin{align*}
R_k \le \frac{1}{\mu}\norm{(I-P_k)\nabla f(y_k)}^2 + \frac{\theta\mu}{4}\norm{y_k-x^*}^2
\le  \frac{\Crel^2}{\mu}\eta^2(X_k,y_k)  + \frac{\theta\mu}{4}\norm{y_k-x^*}^2
\end{align*}
%
If the criterion in step (3) of the algorithm does not hold, we have
%
\begin{align*}
\eta(x_{k+1},X_k) \le \qred \eta(x_{k},X_k) + \lambda(f(x_k)-f(x_{k+1})
\end{align*}
%
Otherwise, we have
%
\begin{align*}
\eta(x_{k+1},X_k) \le \qred \eta(x_{k+1},X_k)\le \qred \eta(x_{k},X_k)+ \Cstab\norm{x_{k+1}-x_{k}}
\end{align*}
%

%




%-----------------------------------------------
\printbibliography
%-----------------------------------------------
%
%-------------------------------------------
\end{document}      
%-------------------------------------------